---
title: "The Ideas.RePEc Listings of Articles in Economics"
author: "Senan Hogan-Hennesssy"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: github_document
---
```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = F, fig.align = 'center')
library(tidyverse)
library(data.table)
library(stringi)
library(ggplot2)
library(RColorBrewer)
library(ggrepel)
library(tidytext)
library(scales)
theme_set(theme_bw())
```

## WORK IN PROGRESS -- to be finished this week

The NBER series is a great source of research in our profession, but it isn't the full picture of research in the field.
It is, after all, a listing of working papers which are by definition not yet publications -- though most do go on to be published,
as I noted in the [last post.](https://github.com/shoganhennessy/Econ_text_data/blob/master/Blog_post_exploration/1_NBER_working_papers.md)

```{r, echo = T}
# Load the data on Economists
Economists.data <- fread('../Data/RePEc_data/Economists_data/Economists_repec_data.csv')
# Load the data on published articles
Articles.data <- fread('../Data/RePEc_data/Papers_data/Journal_articles_repec.csv')
# Load data on Journals, collected 09/01/2020
# https://ideas.repec.org/top/top.journals.all.html
Journals.data <- fread('../Data/RePEc_data/Journals_data/Journal_info_repec.csv')
```
Forunately, [Research Papers in Economics (RePEc)'s Ideas](https://ideas.repec.org/) provides a large listing of economics research which I can use as a complement to the NBER listings as it lists information on `r nrow(Economists.data)` economists 
and `r nrow(Articles.data)` published articles papers -- considering only those in English as of 31 December 2019.

> ''IDEAS is the largest bibliographic database dedicated to Economics and available freely on the Internet. 
> Based on RePEc, it indexes over 3,000,000 items of research, including over 2,700,000 that can be downloaded in full text.''
> 
> -- What is IDEAS?  [Ideas.RePEc.org/](https://ideas.repec.org/) 


### Published Articles

`r nrow(Articles.data)` is a huge number of articles to consider, and as far as I can gather, it is the
largest set yet analysed -- Hamermesh ([2007](https://www.nber.org/papers/w6761),
[2012](https://www.nber.org/papers/w18635)) consider only subsamples of the top three journals for example.
Lucky for us, computing power and open access to research
have advanced enough to allow analysis of a more complete view of research in economics.

```{r, echo=T}
# Get the top Journal
top_journal <- Journals.data %>% 
  arrange(rank) %>% head(1) %>% pull(journal_title)
# Get the bottom Journal
bottom_journal <- Journals.data %>% 
  arrange(rank) %>% tail(1) %>% pull(journal_title)

# SUbset to the top 200 journals
Journals.data <- Journals.data %>% head(200)
```
The set is certainly much larger than the NBER listings, but it is so far unclear whether it is 
representative of the universe of economics publications.
Similarly the universe of economics publications may turn out not to be the target of these data:
a simple scan of the journals included in these data show some *not so important*
publications (also seen at the bottom of [this table](https://ideas.repec.org/top/top.journals.all.html)).
Perhaps economics articles published in *`r top_journal`* are worth keeping, 
yet those in *`r bottom_journal`* are worth taking out of the sample.
SO for convenience, I'm subsetting to the top 200 journals.






So who are the most problific authors, by article count?

...


Most profilific publications?  // Compared to the ranking listing.

...





### Economists

Describe set of data on economists, including PhD attendence.
note: most entries are empty, however.




opportunity link to NBER working paper and measure change from WP to published article

