---
title: "The Ideas.RePEc Listings of Articles in Economics"
author: "Senan Hogan-Hennessy"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: github_document
---
```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = F, fig.align = 'center')
library(tidyverse)
library(data.table)
library(tidytext)
# https://github.com/dgrtwo/fuzzyjoin
library(fuzzyjoin)
```

## WORK IN PROGRESS -- to be finished soon.

The NBER series is a great source of research in our profession, but it isn't the full picture of research in the field.
It is, after all, a listing of working papers which are by definition not yet publications -- though most do go on to be published,
as I noted in the [last post.](https://github.com/shoganhennessy/Econ_text_data/blob/master/Blog_post_exploration/1_NBER_working_papers.md)

```{r, echo = F}
# Load the data on Economists
Economists.data <- fread('../Data/RePEc_data/Economists_data/Economists_repec_data.csv')
# Load the data on published articles
Articles.data <- fread('../Data/RePEc_data/Papers_data/Journal_articles_repec.csv')
# Load data on Journals, collected 09/01/2020
# https://ideas.repec.org/top/top.journals.all.html
Journals.data <- fread('../Data/RePEc_data/Journals_data/Journal_info_repec.csv')
# Load Economist-article individual links -- made externally in Python
Economists_articles.links <- fread('../Data/RePEc_data/Economists_data/Economists_articles_repec.csv')
```
Forunately, [Research Papers in Economics (RePEc)'s Ideas](https://ideas.repec.org/) provides a large listing of economics research which I can use as a complement to the NBER listings as it lists information on `r nrow(Economists.data)` economists and `r nrow(Articles.data)` published articles papers -- considering only those in English as of 9 January 2020.

> ''IDEAS is the largest bibliographic database dedicated to Economics and available freely on the Internet. 
> Based on RePEc, it indexes over 3,000,000 items of research, including over 2,700,000 that can be downloaded in full text.''
> 
> -- What is IDEAS?  [Ideas.RePEc.org/](https://ideas.repec.org/) 

`r nrow(Articles.data)` is a huge number of articles to consider, and comparable in size to articles that use the near complete set of publcations.^[[Angrist et al. 2015](https://www.nber.org/papers/w23698) link Econlit and Web of Science for similar coverage, for example.]
RePEc has certain advantages: it is more open and it has a clear link to authors -- and importantly some of their characteristics -- to working and published papers, all within the field of ecoomics.
It is however, a repository only within the economics sphere: citations are only counted among other papers in the database, and not among the universe of publications.

### Citations

```{r, echo=T}
# The most prolific articles, within these data
Articles.data %>% 
  select(author, title, publication_date, journal_title, citation_count) %>%
  arrange(-citation_count) %>% head(5) %>%
  knitr::kable(format = 'markdown')
```



```{r, echo=T}
# The most prolific authors, within these data
## WHo have the most citations, in these data
Economists_articles.links %>% 
  left_join(select(Economists.data, url, name) %>% 
              rename(author_name = name, author_url = url), by = 'author_url') %>%
  left_join(rename(Articles.data, article_url = url), by = 'article_url') %>%
  select(author_name, title, publication_date, journal_title, citation_count) %>%
  filter(!is.na(citation_count)) %>%
  group_by(author_name) %>%
  summarise(total_citations = sum(citation_count, na.rm = T),
            mean_citations = mean(citation_count, na.rm = T),
            total_articles = n()) %>%
  arrange(-total_citations) %>% head(10) %>%
  knitr::kable(format = 'markdown')
```


### The Top 5

Quote book (name?) about survey for economists :
an article in the top 5 makes a career?

SHow how many articles in the top 10.
```{r, results='asis'}
Articles.data %>% 
  group_by(journal_title) %>%
  summarise(mean_citations = round(mean(citation_count, na.rm = T)),
            citation_count = sum(citation_count, na.rm = T)) %>%
  arrange(-citation_count) %>% head(10) %>% arrange(-mean_citations) %>%
  knitr::kable(format = 'markdown')
```

^ Test the correlation for an article in top 10 making a career.

^ Compare to people who make 6-10 : weak discontinuity, perhaps.

Draw maps of where they're from
and test a correlation: distance to publication home institution and economist base.
^ Start of testing locality or inclusion.


### Link to the NBER

opportunity link to NBER working paper and measure change from WP to published article

**Next up:** Looking at inclusion in the *NBER family*.

***