---
title: ""
author: "Senan Hogan-Hennessy, "
date: "`r format(Sys.time(), '%d %b %Y')`"
output: github_document
---
This is the second of a series of blog posts motivating my analysis and exploration
of a huge set of data on economics publications, sho I hope you enjoy the posts.
The source code is available in [this repo](https://github.com/shoganhennessy/Econ_text_data/),
so feel free to look over and reproduce from there (which is filled mostly with code in *R* and *Python*).
Please contact me if you would like to use my underlying data for another project --
all of the used data are publicly accessible across the internet yet
my collection of them all may constitute sensitive information.

***

```{r setup, include = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T,
                      fig.align = "center")
library(data.table)
library(tidyverse)
library(plm)
library(lintr)
#### Datafile
# Authors and their citations (in year cited)
appointment_repec.data <-
    fread("../Data/RePEc_data/Economists_data/Appointment_ref_repec.csv")
```


# NBER Inclusion, What is it Good For?

> I tried to post a comment on NBER (National Bureau of Economic Research), but was rejected for not being an approved “NBER family member”.
> If any of our readers is a ”NBER family member” feel free to post the above.
> -- Judea pearl on Angrist (2017) http://ideas.repec.org/a/aea/jecper/v31y2017i2p125-44.html
<!-- blog post at http://causality.cs.ucla.edu/blog/index.php/2017/02/ -->

Topic for blog post 3 is discussion of inclusion in nber — see quote above.
Draw some graphs of how NBER members fill the economics publishing sphere,
and establish the issue of benefits from NBER membership.
^ Will require reforming the time series listing of fellow appointments -- easy work yet.


Think about whether sharing of publications/work over the internet leads to more or less inequality -> 
perhaps a better question for the upcoming network analyses.


Look over the paper here:
"Undergraduate Econometrics Instruction: Through Our Classes, Darkly"
Joshua D. Angrist, Joern-Steffen Pischke
http://www.nber.org/papers/w23144

Comment on the same here:
https://p-hunermund.com/2017/02/22/judea-pearl-on-angrist-and-pischke/


```{r, echo = F, fig.height = 5, fig.width = 10}
# find year of appointment and years ref since
appointment_repec.data <- appointment_repec.data %>%
    mutate(appt_year_nber = appt_date_nber %>%
               as.Date(format = "%Y-%m-%d") %>% format("%Y")) %>%
    mutate(appt_since_nber = factor(as.integer(ref_year) - as.integer(appt_year_nber),
                                    levels = c(-1, -15:-2, 0:15)))
# Event linear model
appointment.eventstudy <- appointment_repec.data %>%
    #filter(appt_rank_nber == "RA") %>%
    pdata.frame(index = c("name_repec", "ref_year")) %>%
    plm(log(ref_count) ~ 1 + appt_since_nber, data = .,
        effect = "twoways", model = "within")
# Save note of observation number
anotation_note <- appointment.eventstudy %>% summary() %>% capture.output() %>% nth(7)
# And cluster the SE, keeping relevant coefficients
appointment.eventstudy <- appointment.eventstudy %>%
    lmtest::coeftest(vcov = vcovDC(appointment.eventstudy, type = "sss")) %>%
    broom::tidy() %>%
    filter(str_detect(term, "appt_since_nber")) %>%
    mutate(coefficient = term %>% str_replace("appt_since_nber", "") %>% as.integer(),
           conf.low  = estimate - 1.96*std.error,
           conf.high = estimate + 1.96*std.error)
# Graph the event
appointment.eventstudy %>%
  mutate(estimate_before  = ifelse(coefficient < 0 , estimate, NA),
         estimate_after   = ifelse(0 <= coefficient, estimate, NA),
         conf.low_before  = ifelse(coefficient < 0,  conf.low, NA),
         conf.low_after   = ifelse(0 <= coefficient, conf.low, NA),
         conf.high_before = ifelse(coefficient < 0,  conf.high, NA),
         conf.high_after  = ifelse(0 <= coefficient, conf.high, NA)) %>%
  ggplot(aes(x = coefficient)) +
  geom_point(aes(y = estimate)) +
  stat_smooth(aes(y = estimate), geom = "line", colour = "blue", span = 0.5, alpha = 0.5, se = F) +
  geom_line(aes(y = conf.low_before), linetype = "dashed") +
  geom_line(aes(y = conf.low_after), linetype = "dashed") +
  geom_line(aes(y = conf.high_before), linetype = "dashed") +
  geom_line(aes(y = conf.high_after), linetype = "dashed") +
  scale_x_continuous(name = "Years after appointment", breaks = seq(-15, 15, by = 5)) +
  scale_y_continuous(name = "Coefficient estimate") +
  theme_bw() + theme(plot.title = element_text(hjust = 0.5)) +
  labs(title   = "Log Citations by year to author's NBER appointment",
       caption = anotation_note)
```


### Graph years since PhD



### Link to the NBER

Draw maps of where they"re from
and test a correlation: distance to publication home institution and economist base.
^ Start of testing locality or inclusion.

### Coauthorship

Look at how much more likely to coauthor with other members after being appointed.


```{r postcheck, include=F}
#lintr::lint("Blog_post_exploration/3_NBER_inclusion.Rmd")
```
