---
title: ""
author: "Senan Hogan-Hennessy, "
date: "`r format(Sys.time(), '%d %b %Y')`"
output: github_document
---
This is the second of a series of blog posts motivating my analysis and exploration
of a huge set of data on economics publications, sho I hope you enjoy the posts.
The source code is available in [this repo](https://github.com/shoganhennessy/Econ_text_data/),
so feel free to look over and reproduce from there (which is filled mostly with code in *R* and *Python*).
Please contact me if you would like to use my underlying data for another project --
all of the used data are publicly accessible across the internet yet
my collection of them all may constitute sensitive data.

***

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T,
                      fig.align = "center")
library(data.table)
library(tidyverse)
library(plm)
library(lintr)
#### Datafiles
```

# Network Economics and the Network of Economists

Read paper on connections among congressmen:
"Influencing Connected Legislators," Marco Battaglini and Eleonora Patacchini
https://www.journals.uchicago.edu/doi/10.1086/700074

Also newer on central points in the networks
http://www.nber.org/papers/w27557

R package for network analysis:
https://www.rdocumentation.org/packages/econet/

```{r}
library(econet)
data("a_db_alumni")
```


### Link to the NBER

Draw maps of where they're from
and test a correlation: distance to publication home institution and economist base.
^ Start of testing locality or inclusion.

```{r postcheck, include=F}
#lintr::lint("Blog_post_exploration/4_network_economists.Rmd")
```
